Temporal Dynamics of Normalization Reweighting

### 2022-10-01 manuscript.tex
Added code for automatically downloading data from OSF

We then performed a timecourse analysis, in which we analysed each 11-second trial using a sliding 1-second time window. The top panel of Figure \ref{fig:Pilotdata}c shows the response at the target frequency (7Hz) to a single stimulus of 32\% contrast (black), and the response at 7Hz when the 32\% contrast mask is added (blue). Analogous responses are shown at three other frequencies - the mask frequency (5Hz), and the second harmonics of both target and mask frequencies (14Hz, 10Hz), at which strong responses were also found (see spectra in Figure \ref{fig:Pilotdata}b). The reduction in signal strength when the mask component is added illustrates the masking effect.

@@

{\centering \includegraphics{temp/figures/Fig2}

@@

{\centering \includegraphics{temp/figures/Fig3}

@@

{\centering \includegraphics{temp/figures/Fig4}

@@

To investigate whether normalization reweighting effects differ with respect to autistic traits, we then split each dataset (averaged across temporal frequency) using median AQ score (for the EEG experiments) or according to diagnostic group (autism vs controls) for the MEG data. Figure \ref{fig:AQanalysis}a-c shows distributions of AQ scores for each experiment, and indicates for the pilot and EEG data which participants were in the high (purple) and low (green) AQ groups. The median AQ scores were 14 for the pilot data, and 18 for the EEG data, consistent with previous results for the general population (Baron-Cohen et al., 2001). In the MEG experiment, AQ scores for the autism group (mean 36.1) and the control group (mean 16.7) were significantly different (\emph{t} = 6.00, \emph{df} = 14.2, \emph{p} \textless{} 0.001), with minimal overlap (one participant with an autism diagnosis had an AQ score marginally lower than the highest AQ scores from the control group).

@@

We did not observe clear differences in the timecourse between monocular and dichoptic suppression. This is important, because the dichoptic arrangement bypasses early stages of processing before the cortex (e.g.~the retina and lateral geniculate nucleus). It suggests that the dynamic increases in suppression occur in the cortex, consistent with our MEG data that find evidence of reweighting in V1 (see Fig \ref{fig:MEGdata}), and with previous neurophysiological work (Aschner et al., 2018). It is currently unclear whether these effects originate in V1, or might involve feedback from higher areas. The similarity between monocular and dichoptic effects also differs from work on adaptation to individual mask components. In both physiological (Li et al., 2005; \textbf{Sengpiel2006?}) and psychophysical (Baker et al., 2007) paradigms, adapting to a dichoptic mask reduces its potency, whereas adapting to a monocular mask has little or no effect. Normalization reweighting offers an explanation for why monocular masks presented in isolation do not adapt: if suppressive weights are determined by co-occurrence of stimuli, presentation of an isolated mask will have little effect. However this cannot explain the dichoptic adaptation effects without invoking additional binocular processes, such as competition between summing and differencing channels (e.g. May et al., 2012).

@@

EEG data from Experiments 1 and 2 were first imported into Matlab using components of the EEGlab toolbox (Delorme and Makeig, 2004), and converted into a compressed ASCII format. Primary data analysis was then conducted using a bespoke \emph{R} script. In brief, we epoched each trial and extracted the average timecourse across four occipital electrodes (\emph{Oz}, \emph{POz}, \emph{O1} and \emph{O2}), and then calculated the Fourier transform of this average waveform. We excluded trials for which the Mahalanobis distance of the complex Fourier components exceeded 3 (for details see Baker, 2021). This resulted in 0.25\% of trials being excluded for Experiment 1, 4.51\% of trials for Experiment 2, and 4.51\% of trials for Experiment 3. Next we averaged the waveforms across all remaining trials, and calculated the Fourier transform in a 1-second sliding window to generate timecourses for each participant. We divided the timecourse for the target-only condition by the timecourse for the target + mask condition to produce a suppression ratio. These were then converted to logarithmic (dB) units for averaging, calculation of standard errors, and statistical comparisons. For display purposes we smoothed the timecourses using a cubic spline function, however all statistical comparisons used the unsmoothed data.

For Experiment 3, we performed source localisation using a linearly constrained minimum variance (LCMV) beamformer algorithm, implemented in Brainstorm (Tadel et al., 2011). Structural MRI scans were processed using Freesurfer (Dale et al., 1999) to generate a 3D mesh of the head and brain, and we calculated source weights for each block with reference to a 5-minute empty room recording, usually recorded on the day of the experiment. The matrix of source weights for each block was used in a custom Matlab script to extract signals from V1, identified using the probabilistic maps of Wang et al. (2015). These signals were then imported into R for the main analysis, which was consistent with the EEG analysis described above.

>>>>

We then performed a timecourse analysis, in which we analysed each 11-second trial using a sliding 1-second time window. The top panel of Figure \ref{fig:Pilotdata}c shows the response at the target frequency (7Hz) to a single stimulus of 32\% contrast (black), and the response at 7Hz when the 32\% contrast mask is added (blue). For comparison, a baseline timecourse is also shown (grey), which was the response at 7Hz when a 5Hz stimulus was shown (therefore controlling for attention, blinking etc.). Analogous responses are shown at three other frequencies - the mask frequency (5Hz), and the second harmonics of both target and mask frequencies (14Hz, 10Hz), at which strong responses were also found (see spectra in Figure \ref{fig:Pilotdata}b). The reduction in signal strength when the mask component is added illustrates the masking effect.

@@

{\centering \includegraphics{Figures/EEGdata}

@@

{\centering \includegraphics{Figures/MEGdata}

@@

{\centering \includegraphics{Figures/AQfigure}

@@

To investigate whether normalization reweighting effects differ with respect to autistic traits, we then split each dataset (averaged across temporal frequency) using median AQ score (for the EEG experiments) or according to diagnostic group (autism vs controls) for the MEG data. Figure \ref{fig:AQanalysis}a-c shows distributions of AQ scores for each experiment, and indicates for the pilot and EEG data which participants were in the high (purple) and low (green) AQ groups. The median AQ scores were 14 for the pilot data, and 18 for the EEG data. In the MEG experiment, AQ scores for the autism group (mean 36.1) and the control group (mean 16.7) were significantly different (\emph{t} = 6.00, \emph{df} = 14.2, \emph{p} \textless{} 0.001), with minimal overlap (one participant with an autism diagnosis had an AQ score marginally lower than the highest AQ scores from the control group). These distributions are consistent with previous results for AQ (Baron-Cohen et al., 2001).

@@

We did not observe clear differences in the timecourse between monocular and dichoptic suppression. This is important, because the dichoptic arrangement bypasses early stages of processing before the cortex (e.g.~the retina and lateral geniculate nucleus). It suggests that the dynamic increases in suppression occur in the cortex, consistent with our MEG data that find evidence of reweighting in V1 (see Fig \ref{fig:MEGdata}), and with previous neurophysiological work (Aschner et al., 2018). It is currently unclear whether these effects originate in V1, or might involve feedback from higher areas. The similarity between monocular and dichoptic effects also differs from work on adaptation to individual mask components. In both physiological (Li et al., 2005; Sengpiel and Vorobyov, 2005) and psychophysical (Baker et al., 2007) paradigms, adapting to a dichoptic mask reduces its potency, whereas adapting to a monocular mask has little or no effect. Normalization reweighting offers an explanation for why monocular masks presented in isolation do not adapt: if suppressive weights are determined by co-occurrence of stimuli, presentation of an isolated mask will have little effect. However this cannot explain the dichoptic adaptation effects without invoking additional binocular processes, such as competition between summing and differencing channels (e.g. May et al., 2012).

@@

EEG data from Experiments 1 and 2 were first imported into Matlab using components of the EEGlab toolbox (Delorme and Makeig, 2004), and converted into a compressed ASCII format. Primary data analysis was then conducted using a bespoke \emph{R} script. In brief, we epoched each trial and extracted the average timecourse across four occipital electrodes (\emph{Oz}, \emph{POz}, \emph{O1} and \emph{O2}), and then calculated the Fourier transform of this average waveform. We excluded trials for which the Mahalanobis distance of the complex Fourier components exceeded 3 (for details see Baker, 2021). This resulted in 0.25\% of trials being excluded for Experiment 1, and 4.51\% of trials for Experiment 2. Next we averaged the waveforms across all remaining trials, and calculated the Fourier transform in a 1-second sliding window to generate timecourses for each participant. We divided the timecourse for the target-only condition by the timecourse for the target + mask condition to produce a suppression ratio. These were then converted to logarithmic (dB) units for averaging, calculation of standard errors, and statistical comparisons. For display purposes we smoothed the timecourses using a cubic spline function, however all statistical comparisons used the unsmoothed data.

For Experiment 3, we performed source localisation using a linearly constrained minimum variance (LCMV) beamformer algorithm, implemented in Brainstorm (Tadel et al., 2011). Structural MRI scans were processed using Freesurfer (Dale et al., 1999) to generate a 3D mesh of the head and brain, and we calculated source weights for each block with reference to a 5-minute empty room recording, usually recorded on the day of the experiment. The matrix of source weights for each block was used in a custom Matlab script to extract signals from V1, identified using the probabilistic maps of Wang et al. (2015). These signals were then imported into R for the main analysis, which was consistent with the EEG analysis described above. The outlier rejection procedure excluded 2.47\% of trials for Experiment 3.
